Reduced plasticity and variance in physiological rates for terrestrial ectotherm populations under climate change

### 2024-01-10 ms.qmd
MS updated with Shinichi's comments

Climate change is expected to result in warmer and more variable thermal environments globally. Greater thermal variability is expeced to result in strong selection pressures leading to genetic adaptation and/or the evolution of adaptive phenotypic plasticity. However, such responses depend on the extent of genetic and phenotypic variability, yet most work has focused on mean responses to climate warming ignoring how temperature changes variability. Here, we collect data on marine, freshwater and terrestrial ectotherms to quantify their capacity for acclimation and the extent to which variance in physiological rates change in response to higher temperatures. We use new effect size measures and meta-analysis to understand

@@

where $CV$ is the coefficient of variation defined as $SD / R$.

@@

Overall, $lnRR_{Q_{10_{acclimation}}}$ and $lnRR_{Q_{10_{acute}}}$ differed by only `r mean(avg_change)`% across all habitats (95%CI: `r quantile(avg_change, c(0.025, 0.975))[1]` to `r quantile(avg_change, c(0.025, 0.975))[2]`%). Ectotherms in marine and freshwater environments showed partial compensation of physiological rates ([@fig-1]B) amounting to reduced $lnRR_{Q_{10_{acclimation}}}$ of `r abs(mean(change_f))`% (95% CI: `r quantile(change_f, c(0.025, 0.975))[1]` to `r quantile(change_f, c(0.025, 0.975))[2]`) in freshwater and `r abs(mean(change_m))`% (95% CI: `r quantile(change_m, c(0.025, 0.975))[1]` to `r quantile(change_m, c(0.025, 0.975))[2]`) in marine environments. In contrast, terrestrial ectotherms showed no acclimation (possibly even inverse acclimation) -- showing a `r abs(mean(change_t))`% increase in $lnRR_{Q_{10_{acclimation}}}$ (95% CI: `r quantile(change_t, c(0.025, 0.975))[1]` to `r quantile(change_t, c(0.025, 0.975))[2]`, [@fig-1]B). Considering acute responses of animals acclimated to high temperatures are generally slightly elevated compared to cold acclimated animals (~7%; $\beta$ = `r model1.s1$b[3]`, 95% CI: `r model1.s1$ci.lb[3]` to `r model1.s1$ci.ub[3]`, p `r p_value(model1.s1$pval[3])`), acclimatization is not likely going to provide adaptive benefits under climate change.

@@

Different trait categories showed different accclimation responses across habitat types, however, they mirrored patterns seen overall (see Supplement; @fig-2). Acclimation capacity also did not vary by life-stage given that there were no differences between $lnRR_{Q_{10_{acclimation}}}$ and $lnRR_{Q_{10_{acute}}}$ between adult and juveniles (@fig-3) (Adult-Juvenile (Acute): `r mean(acute_all)`, 95% CI: `r quantile(acute_all, 0.025)[1]` to `r quantile(acute_all, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(acute_all)`; Adult-Juvenile (Acclimation): `r mean(acclim_all)`, 95% CI: `r quantile(acclim_all, 0.025)[1]` to `r quantile(acclim_all, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(acclim_all)`).

@@

p1.2 <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 45, legend.pos = "top.left", trunk.size = 8, branch.size = 2) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +
  geom_segment(aes(x = 2.92, y = 2, xend = 3.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
  geom_segment(aes(x = 1.92, y = 2, xend = 2.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
  geom_segment(aes(x = 0.92, y = 2, xend = 1.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
annotate("text", x = 1, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(model1.2$b[3], 2), "[", round(model1.2$ci.lb[3], 2)," to ", round(model1.2$ci.ub[3], 2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 1.01, y = 4.0, 
         label =paste0(round(mean(change_f), 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
  annotate("text", x = 0.85, y = 3, label = "P < 0.001", hjust = 0, size = 6) + 
  annotate("text", x = 2, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(mean(m_acclim - m_acute), 2), "[", round(quantile(m_acclim - m_acute, 0.025), 2)," to ", round(quantile(m_acclim - m_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 2.01, y = 4.0, 
         label =paste0(round(mean(change_m), 2), "%"), hjust = 0, size = 6) +
  annotate("text", x = 1.85, y = 3, label = "P = 0.01", hjust = 0, size = 6) +
  annotate("text", x = 3, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(mean(t_acclim - t_acute), 2), "[", round(quantile(t_acclim - t_acute, 0.025), 2)," to ", round(quantile(t_acclim - t_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 3.01, y = 4.0, 
         label =paste0(round(mean(change_t), 2), "%"), hjust = 0, size = 6) +
  annotate("text", x = 2.85, y = 3, label = "P = 0.36", hjust = 0, size = 6) + scale_fill_manual(values = c(col)) + ylim(-0.8, 5.05) +  annotate("text", label= paste0("k = ", k_lnrr$k, " (", k_lnrr$sdy, ")"), x = k_lnrr$moderator, y = 5)

  p1.2 <- p1.2 + theme_classic() + theme(axis.title  = element_text(size = 24),
        axis.text = element_text(size = 15))

  pdf("./output/figures/fig1.pdf", useDingbats=FALSE, width = 12, height = 9.6)
    p1.2
  dev.off()
  ggsave(p1.2, filename = "./output/figures/fig1.pdf", width = 12, height = 9.6, device = "pdf")

@@

  p2 <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 8, branch.size = 2) +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + annotate("text", x = 1.3, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(f_cvr_mean), 2), "[", round(quantile(f_cvr_mean, 0.025), 2)," to ", round(quantile(f_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 1.15, y = -1.5, label = paste0("P = ", round(pmcmc(f_cvr_mean), digit = 2)), hjust = 0, size = 6) + 
  annotate("text", x = 1.15, y = 1, 
           label =paste0(round(mean(exp(f_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + 
  annotate("text", x = 2.3, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(m_cvr_mean), 2), "[", round(quantile(m_cvr_mean, 0.025), 2)," to ", round(quantile(m_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 2.15, y = -1.5, label = paste0("P = ", round(pmcmc(m_cvr_mean), digit = 2)), hjust = 0, size = 6) +
  annotate("text", x = 2.15, y = 1, 
           label =paste0(round(mean(exp(m_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + annotate("text", x = 3.4, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 2), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 3.25, y = 1, 
           label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
    annotate("text", x = 3.25, y = -1.5, label = paste0("P = ", round(pmcmc(t_cvr_mean), digit = 3)), hjust = 0, size = 6) + scale_fill_manual(values = c(col)) + ylim(-2, 2) + 
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = c(1.15, 2.15, 3.25), y = 1.8)

  p2 <- p2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 15),
                                       legend.title = element_text(size = 15),
                                        legend.text = element_text(size = 15))

     ggsave(p2, filename = "./output/figures/fig2.png", width = 10, height = 9.5)

```{r, fig-3}
#| label: fig-3
#| fig-cap: "Estimated mean acclimation and acute $lnRR_{Q_{10}}$ for adult and juevenile life-history stages for A) Marine, B) Freshwater and C) Terrestrial ectotherms. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals."

p1 <- orchard_plot(model1.4, mod = "type", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE, legend.pos = "top.left", angle = 45) + labs(title = "Marine") + ylim(c(-2, 2.2))

p2 <- orchard_plot(model1.4, mod = "type", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE)+ labs(title = "Freshwater") + theme(legend.position = "none", axis.text.y = element_blank())

p3 <- orchard_plot(model1.4, mod = "type", at = list(c_accl_time = 0, habitat = "t"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE)+ labs(title = "Terrestrial")+ theme(legend.position = "none", axis.text.y = element_blank())

fig3 <- (p1 | p2 | p3) + plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(fig3, filename = "./output/figures/fig3_lh.png", width = 14, height = 4.5)

fig3 <- magick::image_read("./output/figures/fig3_lh.png")
fig3

@@

```{r, fig-cvrlh}
#| label: fig-cvrlh
#| fig-cap: "Estimated mean $lnCVR_{Q_{10}}$ for adult and juevenile life-history stages for A) Marine, B) Freshwater and C) Terrestrial ectotherms. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals."
#| echo: false
#| warning: false

p1 <- orchard_plot(model2.4, mod = "type", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE, legend.pos = "top.left", angle = 45) + labs(title = "Marine") + ylim(c(-2, 2.2))

p2 <- orchard_plot(model2.4, mod = "type", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE)+ labs(title = "Freshwater") + theme(legend.position = "none", axis.text.y = element_blank())

p3 <- orchard_plot(model2.4, mod = "type", at = list(c_accl_time = 0, habitat = "t"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE)+ labs(title = "Terrestrial")+ theme(legend.position = "none", axis.text.y = element_blank())

fig4 <- (p1 | p2 | p3) + plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(fig4, filename = "./output/figures/fig4_lh.png", width = 14, height = 4.5)

fig4 <- magick::image_read("./output/figures/fig4_lh.png")
fig4
```

```{r, fig-cvrtrait}
#| label: fig-cvrtrait
#| fig-cap: "Acute and Acclimation $lnCVR_{Q_{10}}$ across traits for A) marine, B) freshwater and C) terrestrial systems. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals."
#| echo: false

ptm <- orchard_plot(model2.3m, mod = "trait_category", group = "record_num", weights = "prop",  xlab = TeX("$lnCVR_{Q_{10}}$"), angle = 0, condition.lab = "Type") +
  scale_x_discrete(labels = c("Mito_leak" = "Proton Leak",
                              "Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate",
                              "Max_MR" = "Maximum Metabolic Rate",
                              "Mito_oxidation" = "OXPHOS")) + theme_classic() 
#ptm$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

ptf <- orchard_plot(model2.3f, mod = "trait_category",  group = "record_num", weights = "prop",  xlab = TeX("$lnCVR_{Q_{10}}$"), angle = 0, condition.lab = "Type") +
  scale_x_discrete(labels = c("Mito_leak" = "Proton Leak",
                              "Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate",
                              "Max_MR" = "Maximum Metabolic Rate",
                              "Mito_oxidation" = "OXPHOS")) + theme_classic()
#ptf$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

ptt <- orchard_plot(model2.3t, mod = "trait_category",  group = "record_num", weights = "prop",  xlab = TeX("$lnCVR_{Q_{10}}$"), angle = 0, condition.lab = "Type") +
  scale_x_discrete(labels = c("Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate",
                              "Max_MR" = "Maximum Metabolic Rate")) + theme_classic() + labs(legend = "Type")
#ptt$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

figcvrtrait <- (ptm + ptf + ptt) + plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(figcvrtrait, filename = "./output/figures/figcvrtrait.png", height = 8.5, width = 26)

@@

p8.2 <- bubble_plot(model3.8, mod = "acf_all_c", group = "record_num", by = "habitat",  ylab = TeX("$lnRR_{Q_{10}}$"), xlab = "Thermal Predictability", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text2, aes(x = x, y = y, label = label), size = 5) + 
geom_text(data = dat_text2, aes(x = x, y = y-0.3, label = label2), size = 5) +
geom_text(data = dat_text2, aes(x = x, y = y-0.65, label = p), size = 5) +
theme(legend.position = "bottom", 
      strip.background = element_blank(), 
      strip.text = element_blank(), 
      axis.text = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis),
      legend.text = element_text(size = 16),
      legend.title = element_text(size = 16))

@@

p9.2 <- bubble_plot(model4.8, mod = "acf_all_c", group = "record_num", by = "habitat",  ylab = TeX("$lnCVR_{Q_{10}}$"), xlab = "Thermal Predictability", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text2, aes(x = x, y = y, label = label), size = 5) + 
geom_text(data = dat_text2, aes(x = x, y = y-0.6, label = label2), size = 5) +
geom_text(data = dat_text2, aes(x = x, y = y-1.2, label = p), size = 5) +
theme(legend.position = "bottom", 
      strip.background = element_blank(), 
      strip.text = element_blank(), 
      axis.text = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis),
      legend.text = element_text(size = 16),
      legend.title = element_text(size = 16))

@@

```{r, fig-fig6}
#| label: fig-fig6
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Model predictions for the expected change in acclimation $lnCVR_{Q_{10}}$ across the globe for terrestrial, marine and freshwater ecthotherms. Predictions consider the uncertainty in random effects (i.e., species, phylogeny, study).

@@

Understanding acclimation capacity and how variation in physiological rates change across populations and species is important for predicting the ecological and evolutionary consequences of climate change [@Chevin2017; @Chevin2010; @Bolnick2011]. Here, we show that the relative variance in physiological rates is expected to decrease for ectotherms as temperatures increase, but particularly so for terrestrial ectotherms (~28%). Decreases in the relative variance of physiological rates are larger than the expected benefits from acclimation in marine and terrestrial ectotherms (*marine*: 9% acclimation vs. 15% reduction in variance; *terrestrial*: no acclimation vs 28% reduction in variance). Such effects are expected to result in a decrease in the variability in physiological rates for marine and terrestrial ectotherms between ~4-13% under future climate change projections. Our results provide a more complete picture of the impacts that climate warming will have on ectotherm populations, with important ecological and evolutionary implications that require further empirical and theoretical attention. We discuss our findings in the context of the current literature and highlight key areas for future research.

@@

>>>>

Climate change is expected to result in warmer and more variable thermal environments globally. Greater thermal variability is expeced to result in strong selection pressures leading to genetic adaptation and/or the evolution of adaptive phenotypic plasticity. However, such responses depend on genetic and phenotypic variability, yet most work has focused on plastic changes in mean responses to climate warming ignoring how temperature changes variability. Here, using a large database of physiological rate data on marine, freshwater and terrestrial ectotherms we quantify the capacity for acclimation and the extent to which relative variance in physiological rates change in response to higher temperatures. Using new effect size measures and meta-analysis we show that physiological acclimation and changes in relative variance do not depend on past temperature variability and predictability experienced by a population or differ between life stages. Across habitats there was a general tendency for relative variation in physiological rates to decrease at higher temperatures. Freshwater ectotherms are capable of acclimating and such responses will likely be beneficial considering that these effects are larger than reductions in relative variance in physiological rates. In contrast, acclimation responses of marine organisms coincide with a nearly equal reduction in relative variance in physiolgical rates, suggesting acclimation may not be as beneficial as previously thought. Relative variance reductions are particularly pronounced for terrestrial ectotherms and this coincides with a general inability to acclimate highlighting the vulnerability of terrestrial ectotherms to climate change. Our results highlight the importance of considering changes in variance in physiological rates when predicting the capacity of populations to persist in and adapt to novel environments.

@@

where $CV$ is the coefficient of variation defined as $SD / R$. We refer to $lnCVR_{Q_{10}}$ as relative variance because variance changes are relative to the mean.

@@

Overall, $lnRR_{Q_{10_{acclimation}}}$ decreased from $lnRR_{Q_{10_{acute}}}$ by only `r abs(mean(avg_change))`% across all habitats (95%CI: `r quantile(avg_change, c(0.025, 0.975))[1]` to `r quantile(avg_change, c(0.025, 0.975))[2]`%). Ectotherms in marine and freshwater environments showed partial compensation of physiological rates ([@fig-1]B) amounting to reduced $lnRR_{Q_{10_{acclimation}}}$ of `r abs(mean(change_f))`% (95% CI: `r quantile(change_f, c(0.025, 0.975))[1]` to `r quantile(change_f, c(0.025, 0.975))[2]`) in freshwater and `r abs(mean(change_m))`% (95% CI: `r quantile(change_m, c(0.025, 0.975))[1]` to `r quantile(change_m, c(0.025, 0.975))[2]`) in marine environments. In contrast, terrestrial ectotherms showed no acclimation (possibly even inverse acclimation) -- showing a `r abs(mean(change_t))`% increase in $lnRR_{Q_{10_{acclimation}}}$ (95% CI: `r quantile(change_t, c(0.025, 0.975))[1]` to `r quantile(change_t, c(0.025, 0.975))[2]`, [@fig-1]B). Considering acute responses of animals acclimated to high temperatures are generally slightly elevated compared to cold acclimated animals (~7%; $\beta$ = `r model1.s1$b[3]`, 95% CI: `r model1.s1$ci.lb[3]` to `r model1.s1$ci.ub[3]`, p `r p_value(model1.s1$pval[3])`), acclimatization is not likely going to provide adaptive benefits under climate change.

@@

Different trait categories showed different accclimation responses across habitat types, however, they mirrored patterns seen overall (see Supplement; @fig-2). Acclimation capacity also did not vary by life-stage given that there were no differences between $lnRR_{Q_{10_{acclimation}}}$ and $lnRR_{Q_{10_{acute}}}$ between adult and juveniles (@fig-cvrlh A-C) (Adult-Juvenile (Acute): `r mean(acute_all)`, 95% CI: `r quantile(acute_all, 0.025)[1]` to `r quantile(acute_all, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(acute_all)`; Adult-Juvenile (Acclimation): `r mean(acclim_all)`, 95% CI: `r quantile(acclim_all, 0.025)[1]` to `r quantile(acclim_all, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(acclim_all)`).

@@

p1.2 <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 45, legend.pos = "top.left", trunk.size = 8, branch.size = 2) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +

  annotate("text", x = 1.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(model1.2$b[3], 2), "[", round(model1.2$ci.lb[3], 2)," to ", round(model1.2$ci.ub[3], 2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 1.20, y = 1.4, 
         label =paste0(round(mean(change_f), 3), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
  annotate("text", x = 1.20, y = -0.8, label = "P < 0.001", hjust = 0, size = 6) +

  annotate("text", x = 2.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(mean(m_acclim - m_acute), 2), "[", round(quantile(m_acclim - m_acute, 0.025), 2)," to ", round(quantile(m_acclim - m_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 2.20, y = 1.4, 
         label =paste0(round(mean(change_m), 3), "%"), hjust = 0, size = 6) +
  annotate("text", x = 2.20, y = -0.8, label = "P = 0.01", hjust = 0, size = 6) +

  annotate("text", x = 3.5, y = 0, 
         label =TeX(paste0("$\\beta = $", round(mean(t_acclim - t_acute), 2), "[", round(quantile(t_acclim - t_acute, 0.025), 2)," to ", round(quantile(t_acclim - t_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 3.20, y = 1.4, 
         label =paste0(round(mean(change_t), 3), "%"), hjust = 0, size = 6) +
  annotate("text", x = 3.20, y = -0.8, label = "P = 0.36", hjust = 0, size = 6) +

  scale_fill_manual(values = c(col)) + ylim(-0.8, 2) +  annotate("text", label= paste0("k = ", k_lnrr$k, " (", k_lnrr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.7, size = 5)

  p1.2 <- p1.2 + theme_classic() + theme(axis.title  = element_text(size = 24),
        axis.text = element_text(size = 15), legend.title = element_text(size = 15),
                               legend.text = element_text(size = 15),
                               title = element_text(size = 24))

  ggsave(p1.2, filename = "./output/figures/fig1.png", width = 9.1, height = 9.2, device = "png")

@@

  x1 = 1.4
  x2 = 2.4
  x3 = 3.4

  p2 <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 8, branch.size = 2, k.pos = "none") +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +

  annotate("text", x = x1, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(f_cvr_mean), 3), "[", round(quantile(f_cvr_mean, 0.025), 2)," to ", round(quantile(f_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x1-0.15, y = -1.5, label = paste0("P = ", round(pmcmc(f_cvr_mean), digit = 2)), hjust = 0, size = 6) + 
  annotate("text", x = x1-0.15, y = 1, 
           label =paste0(round(mean(exp(f_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) +

  annotate("text", x = x2, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(m_cvr_mean), 3), "[", round(quantile(m_cvr_mean, 0.025), 2)," to ", round(quantile(m_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x2-0.15, y = -1.5, label = paste0("P = ", round(pmcmc(m_cvr_mean), digit = 2)), hjust = 0, size = 6) +
  annotate("text", x = x2-0.15, y = 1, 
           label =paste0(round(mean(exp(m_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) +

  annotate("text", x = x3, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 3), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x3-0.15, y = 1, 
           label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
    annotate("text", x = x3-0.15, y = -1.5, label = paste0("P = ", round(pmcmc(t_cvr_mean), digit = 3)), hjust = 0, size = 6) + scale_fill_manual(values = c(col)) + ylim(-2, 2) + 
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.3, size = 5)

  p2 <- p2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 15),
                                       legend.title = element_text(size = 15),
                               legend.text = element_text(size = 15),
                               title = element_text(size = 24))

  ggsave(p2, filename = "./output/figures/fig2.png", width = 9.3, height = 8.8)

@@

```{r, fig-cvrlh}
#| label: fig-cvrlh
#| fig-cap: "Estimated mean acclimation and acute $lnRR_{Q_{10}}$ (A-C) and $lnCVR_{Q_{10}}$ (D-F) for adult and juevenile life-history stages for Marine (A & D), Freshwater (B & E) and Terrestrial (C & F) ectotherms. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals."
#| echo: false
#| warning: false

add_theme  <- function(){theme(axis.title  = element_text(size = 24),
                               axis.text.y = element_text(size = 15),
                               axis.text.x = element_text(size = 15),
                              legend.title = element_text(size = 15),
                               legend.text = element_text(size = 15),
                               title = element_text(size = 24))}
branch.size = 2
trunk.size = 5

p1 <- orchard_plot(model1.4, mod = "type", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE, legend.pos = "top.left", angle = 45, branch.size = branch.size, trunk.size = trunk.size) + labs(title = "Marine") + add_theme() + ylim(-1, 1.8)

p2 <- orchard_plot(model1.4, mod = "type", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45,)+ labs(title = "Freshwater")  + add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) + ylim(-1, 1.8)

p3 <- orchard_plot(model1.4, mod = "type", at = list(c_accl_time = 0, habitat = "t"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Terrestrial") + add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) +  ylim(-1, 1.8)

p1_cvr <- orchard_plot(model2.4, mod = "type", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE, legend.pos = "top.left", angle = 45, branch.size = branch.size, trunk.size = trunk.size) + ylim(c(-2, 2.2)) + add_theme() + ylim(-2.5, 2.5)

p2_cvr <- orchard_plot(model2.4, mod = "type", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)  + add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) + ylim(-2.5, 2.5)

p3_cvr <- orchard_plot(model2.4, mod = "type", at = list(c_accl_time = 0, habitat = "t"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Life Stage", by = "life.hist",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)+ add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) + ylim(-2.5, 2.5)

fig4 <-(p1 | p2 | p3) / (p1_cvr | p2_cvr | p3_cvr) + plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(fig4, filename = "./output/figures/fig4_lh.pdf", width = 19, height = 12, device = "pdf")

fig4 <- magick::image_read("./output/figures/fig4_lh.pdf")
fig4
```

@@

p8.2 <- bubble_plot(model3.8, mod = "acf_all_c", group = "record_num", by = "habitat",  ylab = TeX("$lnRR_{Q_{10}}$"), xlab = "Thermal Predictability", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text2, aes(x = x+1, y = y, label =label), size = 5) + 
geom_text(data = dat_text2, aes(x = x+1, y = y-0.3, label =label2), size = 5) +
geom_text(data = dat_text2, aes(x = x+1, y = y-0.65, label = p), size = 5) +
theme(legend.position = "bottom", 
      strip.background = element_blank(), 
      strip.text = element_blank(), 
      axis.text = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis),
      legend.text = element_text(size = 16),
      legend.title = element_text(size = 16))

@@

p9.2 <- bubble_plot(model4.8, mod = "acf_all_c", group = "record_num", by = "habitat",  ylab = TeX("$lnCVR_{Q_{10}}$"), xlab = "Thermal Predictability", condition.nrow = 1, k = FALSE) + 
theme_classic() + 
scale_fill_manual(values = c(col)) + 
geom_text(data = dat_text2, aes(x = x+1, y = y, label = label), size = 5) + 
geom_text(data = dat_text2, aes(x = x+1, y = y-0.6, label = label2), size = 5) +
geom_text(data = dat_text2, aes(x = x+1, y = y-1.2, label = p), size = 5) +
theme(legend.position = "bottom", 
      strip.background = element_blank(), 
      strip.text = element_blank(), 
      axis.text = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis),
      legend.text = element_text(size = 16),
      legend.title = element_text(size = 16))

@@

```{r, fig-fig6}
#| label: fig-fig6
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Model predictions for the expected change in acclimation $lnCVR_{Q_{10}}$ across the globe for terrestrial, marine and freshwater ecthotherms. Predictions consider the uncertainty in random effects (i.e., species, phylogeny, study).  Predicted change in physiological rate variance (relative variance) for each population based on current temperatures (average from 2018-2022; A-C) as well as the expected change from current temperatures based on future temperature predictions (average from 2096-2100, D). Future climate predictions are the reduction in relative variance expected under a RCP8.5 climate scenerio relative to current climate conditions (% change).

@@

Understanding acclimation capacity and how variation in physiological rates change across populations and species is important for predicting the ecological and evolutionary consequences of climate change [@Chevin2017; @Chevin2010; @Bolnick2011]. Here, we show that the relative variance in physiological rates is expected to decrease for ectotherms as temperatures increase, but particularly so for terrestrial ectotherms (~28%). Decreases in the relative variance of physiological rates are equal to or larger than the expected benefits from acclimation in marine and terrestrial ectotherms (*marine*: 16% acclimation vs. 15% reduction in relative variance; *terrestrial*: no acclimation vs 28% reduction in relative variance; @fig-1). Such effects are expected to result in a decrease in the variability in physiological rates for marine and terrestrial ectotherms between ~4-13% under future climate change projections. Our results provide a more complete picture of the impacts that climate warming will have on ectotherm populations, with important ecological and evolutionary implications that require further empirical and theoretical attention. We discuss our findings in the context of the current literature and highlight key areas for future research.

@@

```{r, fig-cvrtrait}
#| label: fig-cvrtrait
#| fig-cap: "Acute and Acclimation $lnCVR_{Q_{10}}$ across traits for A) marine, B) freshwater and C) terrestrial systems. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. The x-axis is truncated for ease of visualisation."
#| echo: false

k_marinecv=manual_extract(model2.3m, mod = "trait_category", group = "record_num", exclude = "Antiox")
k_freshcv=manual_extract(model2.3f, mod = "trait_category", group = "record_num")
k_terrscv=manual_extract(model2.3t, mod = "trait_category", group = "record_num")
